---
title: "L4-phiho267-zijfe244"
author: "phiho267 & zijfe244"
date: "2019/5/23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, out.width = "320px")
```

## (a) Write an AR(1)-process function in R 
Write a in R 
that simulates data from the AR(1)-process 
$$x_t = \mu +\phi(x_{t-1}-\mu)+\varepsilon_t,\\ \overset{\text{iid}}{\sim} \mathcal{N}(0,\sigma^2) $$
for given values of $\mu$, $\phi$ and $\sigma^2$. Start the process at $x1 =  \mu$ and then simulate
values for $x_t$ for $t = 2, 3,..., T$ and return the vector $x_{1:T}$ containing all time
points. Use $\mu = 10$; $\sigma^2 = 2$ and $T = 200$ and look at some different realizations
(simulations) of x1:T for values of $\phi$ between -1 and 1 (this is the interval of
$\phi$ where the AR(1)-process is stable). Include a plot of at least one realization
in the report. What effect does the value of $\phi$ have on $x_{1:T}$ ?

```{r}
# setting
library(rstan)
library(latex2exp)
library(gridExtra)
library(ggplot2)

rm(list = ls())
set.seed(123456)
```


```{r}
# AR(1) process
AR1 <- function(phi, n=200, mu=10, sigma2=2){
  x <- c()
  x[1] <- mu
  for (i in 2:n) {
    x[i] <- mu + phi*(x[i-1]-mu) + rnorm(1, 0, sqrt(sigma2))
  }
  x
}
```

```{r, echo=FALSE}
iter <- 1:200
phis <- c(-0.9, 0, 0.9)
plot.a <- lapply(phis, function(phi){
  x <- AR1(phi = phi)
  title <- paste0("phi = ",phi)
  data <- data.frame(iter = iter, x=x)
  ggplot(data) +
    geom_line(aes(x=iter, y=x)) +
    ggtitle(title)
})

plot(arrangeGrob(grobs = plot.a))

```

Based on the formula of AR(1)-process, $\phi$ is the parameter representing the impact from previous iterations. when $\phi=0$ all the terms of seem to be independent mutually. On the other hand, there seems to be a positive correlation among simulated terms when $\phi>0$ and negative correlation when $\phi<0$.

## (b) Simulate two AR(1)-processes

Use your function from a) to simulate two AR(1)-processes $x_{1:T}$ with $\phi = 0,3$ and $y_{1,T}$ with $\phi = 0:95$. Now, treat the values of $\mu$, $\phi$ and $\sigma^2$ as unknown and estimate them using MCMC. Implement Stan-code that samples from the
posterior of the three parameters, using suitable non-informative priors of your
choice. [Hint: Look at the time-series models examples in the Stan reference
manual, and note the different parameterization used here.]

 - i. Report the posterior mean, 95% credible intervals and the number of effective
posterior samples for the three inferred parameters for each of the
simulated AR(1)-process. Are you able to estimate the true values?

 - ii. For each of the two data sets, evaluate the convergence of the samplers
and plot the joint posterior of $\mu$ and $\phi$. Comments?

We assume $\mu\sim \mathcal N (0,100)$ and $\sigma^2\sim Inv\  \chi^2(100, 2)$$

```{r echo=FALSE}
set.seed(1111)

x <- AR1(phi=0.3)
y <- AR1(phi=0.95)
N = 200

StanModel='
data {
  int<lower=0> N; 
  vector[N] x;
}
parameters {
  real mu;
  real<lower=0> sigma2;
  real<lower=-1, upper=1> phi;
}
model {
  \\ prior
  mu ~ normal(0,100);
  sigma2 ~ scaled_inv_chi_square(100,2);
  
  \\ model
  x[2:N] ~ normal( mu + phi*(x[1:(N-1)]-mu), sqrt(sigma2));
}'

fit1 <- stan(
  model_code = StanModel,  # Stan program
  data = list(N=N, x=x),    # named list of data
  chains = 1,
  warmup = 1000,          # number of warmup iterations per chain
  iter = 2000,            # total number of iterations per chain
  refresh = 0
)

fit2 <- stan(
  model_code = StanModel,  # Stan program
  data = list(N=N, x=y),    # named list of data
  chains = 1,
  warmup = 1000,          # number of warmup iterations per chain
  iter = 2000,            # total number of iterations per chain
  refresh = 0
)
```

```{r, echo=FALSE}
fit1
re1 <- extract(fit1)
fit2
re2 <- extract(fit2)
```

The mean's in the results for all parameters are quite close to our defined values (besides $\mu$ for $\phi=0.95$). But the 95% credible intervals and number of effective posterior samples are different when $\phi$ are different. The result with $\phi=0.3$ has smaller credible intervals and more effective samples. Anyway, it is okay to estimate sigma2 and phi by `stan`.

```{r echo=FALSE}
grid.arrange(plot(fit1), plot(fit2), ncol = 1)
```

```{r echo=FALSE}
trace1 <- traceplot(fit1, pars = c("mu", "phi"), inc_warmup = TRUE, nrow = 2)
trace2 <- traceplot(fit2, pars = c("mu", "phi"), inc_warmup = TRUE, nrow = 2)
grid.arrange(arrangeGrob(trace1, bottom = "0.3"), arrangeGrob(trace1, bottom = "0.95"), ncol = 2)
```

All parameters converge to the actual values before 300 iterations. However, since the original *y* is a positive correlated process, a continuing increasing/decreasing instance always results in the explosion of estimation for $\mu$. Such phenomena could make the estimation for $\mu$ problematic. Even if $mu$ does not explode, its 95% credible interval would be larger compared with the one with original $\phi=0.3$.

## (c) Cases of campylobacter infections

The data *campy.dat* contain the number of cases of campylobacter infections
in the north of the province Quebec (Canada) in four week intervals from
January 1990 to the end of October 2000. It has 13 observations per year and
140 observations in total. Assume that the number of infections ct at each
time point follows an independent Poisson distribution when conditioned on a
latent AR(1)-process $x_t$, that is
$$ c_t|x_t\sim \mathcal Poisson (\exp (x_t)), $$
where $x_t$ is an AR(1)-process as in a). Implement and estimate the model in
Stan, using suitable priors of your choice. Produce a plot that contains both the
data and the posterior mean and 95% credible intervals for the latent intensity
$\theta_t = \exp (x_t)$ over time. [Hint: Should xt be seen as data or parameters?]

```{r echo=FALSE}
data.c <- read.csv("campy.dat", header = TRUE)
N <- dim(data.c)[1]
```

```{r echo=FALSE}
set.seed(1111)
StanModel.c='
data {
  int<lower=0> N;
  int c[N];
}
parameters {
  real mu;
  real<lower=0> sigma2;
  real phi;
  real x[N];
}
transformed parameters {
  real theta[N];
  theta = exp(x);
}
model {
  // prior
  mu ~ normal(0,50);
  sigma2 ~ scaled_inv_chi_square(100,2);
  
  // model
  for (n in 2:N){
    x[n] ~ normal(mu+phi*(x[n-1]-mu),sqrt(sigma2));
    c[n] ~ poisson(theta[n]);
  }
}
'

set.seed(1111)
datapoi=list(N=N,c=data.c$c)
fit.c<-stan(model_code=StanModel.c,
           data=datapoi,
           warmup=1000,
           iter=2000,
           chains=1,
           refresh=0
           )

re.c <- extract(fit.c)
```

```{r, echo=FALSE}
index <- 1:N
plot(x=index, y=data.c$c, col="black",type="l",ylim=c(0,70))

theta <- exp(colMeans(re.c$x))
lines(x=index, y=theta, col="red", lty=3)

q <- matrix(nrow=2, ncol=140)
for (i in index) {
  q[,i] <- quantile(re.c$x[,i],probs = c(0.025, 0.975))
}
lines(x=index, y=exp(q[1,]), lty=2, col = "pink")
lines(x=index, y=exp(q[2,]), lty=2, col = "pink")
legend(x=1,y=50, legend=c("c","theta"), col=c("black", "red"),lty=1,  cex=1)
```

## (d) Change the prior for $\sigma^2$
Now, assume that we have a prior belief that the true underlying intensity $\theta_t$
varies more smoothly than the data suggests. Change the prior for $\sigma^2$ so that
it becomes informative about that the AR(1)-process increments $\varepsilon_t$ should be
small. Re-estimate the model using Stan with the new prior and produce the
same plot as in c). Has the posterior for $\theta_t$ changed?

Change $\sigma^2$ from $Inv\  \chi^2(100, 2)$ to $Inv\  \chi^2(50, 0.5)$

```{r echo=FALSE}
set.seed(1111)
StanModel.c='
data {
  int<lower=0> N;
  int c[N];
}
parameters {
  real mu;
  real<lower=0> sigma2;
  real phi;
  real x[N];
}
transformed parameters {
  real theta[N];
  theta = exp(x);
}
model {
  // prior
  mu ~ normal(0,50);
  sigma2 ~ scaled_inv_chi_square(50,0.5);
  
  // model
  for (n in 2:N){
    x[n] ~ normal(mu+phi*(x[n-1]-mu),sqrt(sigma2));
    c[n] ~ poisson(theta[n]);
  }
}
'

set.seed(1111)
datapoi=list(N=N,c=data.c$c)
fit.c<-stan(model_code=StanModel.c,
           data=datapoi,
           warmup=1000,
           iter=2000,
           chains=1,
           refresh=0
           )

re.c <- extract(fit.c)

index <- 1:N
plot(x=index, y=data.c$c, col="black",type="l",ylim=c(0,70))

theta <- exp(colMeans(re.c$x))
lines(x=index, y=theta, col="red", lty=3)

q <- matrix(nrow=2, ncol=140)
for (i in index) {
  q[,i] <- quantile(re.c$x[,i],probs = c(0.025, 0.975))
}
lines(x=index, y=exp(q[1,]), lty=2, col = "pink")
lines(x=index, y=exp(q[2,]), lty=2, col = "pink")
legend(x=1,y=50, legend=c("c","theta"), col=c("black", "red"),lty=1,  cex=1)
```

By changing the parameters of prior distribution for $\sigma^2$, true underlying intensity $\theta_t$ is smoother a little bit. However, It is obvious that the 95% credible intervals are smoother compared with $\sigma^2$ with big-variance prior distribution, which is reasonable. 


\newpage
# Lecture code


\newpage
# Appendix

```{r ref.label=knitr::all_labels(), echo = T, eval = F}
```